Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-08-16 ms.tex
Started introduction with some history.  Added appropriate refs.

Radiative transfer is ubiquitous in many astrophysical problems,
encompassing topics such as stellar atmospheres, the interstellar
medium (ISM), galaxy formation, and cosmology.  Thus it is a
well-studied problem \citep[e.g.][and more!]{Rybicki}; however, its
treatment in multi-dimensional calculations is difficult because of
the dependence on seven variables --- three spatial, two angular,
frequency, and time.  The non-local nature of the thermal and
hydrodynamical response to radiation sources further adds to the
difficulty.  In the general case, one must consider all radiation
sources at every point in calculation.  (add: simplify by symmetries
and special cases)

@@

We compare nearly identical simulations but one with radiation
pressure and one without radiation pressure to quantify its effects.
Radial profiles of density, temperature, neutral fraction, and radial
velocity are shown in Figure \ref{fig:rp_profiles} for both
simulations.  Without radiation pressure, the evolution of the
\ion{H}{2} is matches the analytical expectations described in
\S\ref{sec:test5}.  At $t=140$ kyr, the ionization front radius is
increased by $\sim5\% = 0.16$ pc.  However radiation pressure impacts
the system the greatest inside the ionization front.  At $t = 40$ kyr,
the central density is smaller by a factor of 20 with radiation
pressure, but the temperatures are almost equal.  A rarefaction wave
thus propagates toward the center, depicted by the negative radial
velocities at $t = 60$ kyr.  This raises the central density to
$10^{-21}$ g \cubecm~at $t = 80$ kyr.  Afterwards, the radiation
continues to force gas outwards.  From $t = 100$ kyr to $t = 140$ kyr,
the maximum radial velocity of the ionized gas increases from 10
\kms~to 50 \kms.  This leaves behind an even more diffuse medium,
lowering the gas density by a factor of 10 at $t = 140$ kyr.  Thus the
recombination rates are lower, resulting in increased ionization
fractions and temperatures in the \ion{H}{2} region.

@@

Here we test for any inaccuracies in the case of multiple sources.  We
use the same test problem as \citet[][\S5.1.2]{Petkova09}, which has
two sources with luminosities of $5 \times 10^{48}$ ph s$^{-1}$ and
are separated by 8 kpc.  The ambient medium is static with a uniform
density of $10^{-3}$ \cubecm and $T = 10^4$ K.  This setup is similar
to Test 1.  The domain has a resolution of $128^3$ and is 20 kpc in
width, and the problem is run for 500 Myr.

>>>>

Radiative transfer is ubiquitous in many astrophysical problems,
encompassing topics such as stellar atmospheres, the interstellar
medium (ISM), star formation, galaxy formation, supernovae (SNe) and
cosmology.  Thus it is a well-studied problem
\citep[e.g.][]{Mathews65, Rybicki, Mihalas84, Yorke86}; however, its
treatment in multi-dimensional calculations is difficult because of
the dependence on seven variables --- three spatial, two angular,
frequency, and time.  The non-local nature of the thermal and
hydrodynamical response to radiation sources further adds to the
difficulty.  In the general case, one must consider all radiation
sources at every point in calculation.  Fortunately problems can be
simplified by exploiting symmetries and special cases.  The most basic
case was first considered by \citet{Stroemgren39} for a radiation
source ionizing a static uniform neutral medium.  He found that the
delineation between the neutral and ionized medium was found to be
approximately the mean free path of the ionizing radiation.  His
seminar work was expanded upon by \citet{Spitzer48, Spitzer49,
  Spitzer54} and \citet{Spitzer50}, who showed that the ionizing
radiation heated the medium to $T \sim 10^4$ K.  If the density is
equal on both sides of the ionization front, then this
over-pressurized region would expand and drive a shock outwards
\citep[e.g.][]{Oort54, Schatzman55}.  These early works provided the
basis for the modern topic of radiation hydrodynamics.  A decade
later, radiation hydrodynamical numerical models of \ion{H}{2} regions
in spherical symmetry were developed that described the expansion of
the ionization front and the evolution of its associated shock wave
that carries most of the gas outwards \citep[e.g.][]{Mathews65,
  Lasker66, Hjellming66}.  At the same time, theoretical models of
ionization fronts matured and were classified by \citet{Kahn54} and
\citet{Axford61} as either R-type (rare) or D-type (dense).  R-type
fronts travel supersonically with respect to the neutral gas, whereas
D-type front are subsonic.  Furthermore ``weak'' and ``strong'' R-type
fronts move subsonically and supersonically with respect to the
ionized gas.  The same terminology conversely applies to D-type
fronts.

@@

We compare nearly identical simulations but one with radiation
pressure and one without radiation pressure to quantify its effects.
Radial profiles of density, temperature, neutral fraction, and radial
velocity are shown in Figure \ref{fig:rp_profiles} for both
simulations at several times.  Without radiation pressure, the
evolution of the \ion{H}{2} is matches the analytical expectations
described in \S\ref{sec:test5}.  At $t=140$ kyr, the ionization front
radius is increased by $\sim5\% = 0.16$ pc.  However radiation
pressure impacts the system the greatest inside the ionization front.
At $t = 40$ kyr, the central density is smaller by a factor of 20 with
radiation pressure, but the temperatures are almost equal.  A
rarefaction wave thus propagates toward the center, depicted by the
negative radial velocities at $t = 60$ kyr.  This raises the central
density to $10^{-21}$ g \cubecm~at $t = 80$ kyr.  Afterwards, the
radiation continues to force gas outwards.  From $t = 100$ kyr to $t =
140$ kyr, the maximum radial velocity of the ionized gas increases
from 10 \kms~to 50 \kms.  This leaves behind an even more diffuse
medium, lowering the gas density by a factor of 10 at $t = 140$ kyr.
Thus the recombination rates are lower, resulting in increased
ionization fractions and temperatures in the \ion{H}{2} region.

@@

Here we test for any inaccuracies in the case of multiple sources.  We
use the same test problem as \citet[][\S5.1.2]{Petkova09}, which has
two sources with luminosities of $5 \times 10^{48}$ ph s$^{-1}$ and
are separated by 8 kpc.  The ambient medium is static with a uniform
density of $10^{-3}$ \cubecm and $T = 10^4$ K.  This setup is similar
to Test 1.  The domain has a resolution of $128 \times 64 \times 64$
It is 20 kpc in width and is 10 kpc in height and depth.  The problem
is run for 500 Myr.
